---
title: "Weighted data survey tables in R"
author: "John Johnson"
date: "3/9/2020"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`pollster` is an R package for making topline and crosstab tables of simple weighted survey data. The package is designed for use with labelled data, like what you might use the `haven` package to import from Stata or SPSS. It follows tidyverse programming conventions, and output tables are also in the form of a tidy data frame, or tibble.

Only simple weights are currently supported. For complex survey designs, we recommend the excellent [`survey` package](http://r-survey.r-forge.r-project.org/survey/). 

The core functions are:

* `topline()`
* `crosstab()`
* `crosstab_3way()`

Each function also has a twin version which includes a column for the margin of error calculated to include the design effect of the weights.

* `moe_topline()`
* `moe_crosstab()`
* `moe_crosstab_3way()`

There are also two special functions which calculate the design effect component of the margin of error for each survey wave independently.

* `moe_wave_crosstab()`
* `moe_wave_crosstab_3way()`

## Installation

Install it this way.

```
remotes::install_github("jdjohn215/pollster")
```

## Basic useage

`pollster` includes a dataset of Illinois responses to the Current Population Survey's voter registration supplement.

```{r}
library(pollster)
head(illinois)
```

Make a topline table like this. The output is a tibble.

```{r}
topline(variable = maritalstatus, df = illinois, weight = weight)
```

Make a crosstab like this.

```{r}
crosstab(x = educ6, y = maritalstatus, df = illinois, weight = weight)
```

If you prefer, you can also get the output in long format.

```{r}
crosstab(x = educ6, y = maritalstatus, df = illinois, weight = weight, format = "long")
```

A three-way crosstab is just a normal crosstab with a third control variable. Often, this third variable is time.

```{r}
crosstab_3way(x = educ6, y = maritalstatus, z = year, df = illinois, weight = weight)
```

## Making tables and graphs

Wide format is best for displaying table output. Long format is best for making graphs. `pollster` outputs dovetail seamlessly with [`knitr::kable()`](https://www.rdocumentation.org/packages/knitr/versions/1.28/topics/kable) and [`ggplot2::ggplot()`](https://ggplot2.tidyverse.org/). These examples show very basic html table output, but you can customize the appearance of your tables almost endlessly in either html or pdf formats using Hao Zhu's excellent [`kableExtra` package](https://haozhu233.github.io/kableExtra/).

```{r}
library(dplyr)
crosstab(x = sex, y = educ6, df = illinois, weight = weight) %>%
  knitr::kable(digits = 0)
```


```{r}
library(ggplot2)
crosstab(x = sex, y = educ6, df = illinois, weight = weight, format = "long") %>%
  ggplot(aes(educ6, pct, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")
```

Three-way crosstabs are ideal for plotting time series graphs and/or faceted plots.

```{r}
crosstab_3way(x = sex, y = educ6, z = year, df = illinois, weight = weight, format = "long") %>%
  mutate(year = as.numeric(as.character(year))) %>%
  ggplot(aes(year, pct, col = sex)) +
  geom_line() +
  facet_wrap(facets = vars(educ6))
```

## Margin of error

Each `pollster` function comes with a twin function which includes a margin of error column. For example:

```{r}
moe_topline(variable = voter, df = illinois, weight = weight)
```


By default, `moe_crosstab` output comes in long format, but you can also specify wide format.

```{r}
moe_crosstab(x = raceethnic, y = voter, df = illinois, weight = weight, format = "wide")
```

```{r}
moe_crosstab(x = raceethnic, y = voter, 
             df = illinois, 
             weight = weight) %>%
  ggplot(aes(x = pct, y = raceethnic, xmin = (pct - moe), xmax = (pct + moe), color = voter)) +
  geom_pointrange(position = position_dodge(width = 0.2))
```
